---
title: "Example of main text"
output:
  pdf_document:
    toc: true
    toc_depth: 2
  html_document:
    default
header-includes:
   - \usepackage{float}
---

\newpage

\begin{centering}

\vspace{3 cm}

\Huge

{\bf Thesis title}

\vspace{3 cm}

\Large
Author Name

\vspace{3 cm}


\normalsize
Submitted in partial fulfilment of the requirements of the degree of ...

Month Year

\vspace{3 cm}

\normalsize
School of ...

\normalsize
University

\end{centering}

\newpage

# Abstract

"Abstract is an abstract concept."






(The title should always be centred. This has not been done since it's only an example.)


\newpage

# Acknowledgements

Thank you for doing this tutorial!

I hope you learnt something useful from it.

If you have any questions, you can contact me at: `s1650956@sms.ed.ac.uk` (Anna)

\newpage


# RQ and figures

Research Question: are Norway Puffin population trends affected by climate change?

```{r include=F}
# Loading necessary packages
library(tidyverse)  # data transformation
library(reshape2) # restructure and aggregate data
library(ggplot2) # create beautiful graphs

# Load Living planet data
LPI <- read_csv("./data/LPIdata_Feb2016.csv")

# Load climate data
climate_data <- read_csv("./data/Lerwick_temp_data.csv")

# Puffins data transformation
LPI <- LPI %>%
  gather(key = "year", value = "population", select = 26:70) %>%
  mutate(binomial_id = paste(Genus, Species, id, sep = "_")) %>%
  filter(population >= 0) %>%
  group_by(binomial_id) %>%
  mutate(max_year = max(year), min_year = min(year),
         lengthyear = as.integer(max_year) - as.integer(min_year),
         pop_trend = (population - min(population))/(max(population) - min(population))) %>%
  ungroup()

# European puffins only

eu_puffins <- filter(LPI, Genus == "Fratercula", Species == "arctica",
         `Country list` !="Russian Federation") %>%
  select(`Country list`, year, pop_trend, id)  %>%
  group_by(id) %>%
  filter(length(year)>10)  # studies longer than 10 years, it results they have
                            # been carried out exclusively in Norway.

write_csv(eu_puffins, file.path("puffin_toplot.csv"))


# Mean T

mean_t_data <- climate_data %>%
  group_by(year) %>%
  summarise(mean_tmax = mean(tmax),  
            mean_tmin = mean(tmin))

# Merging EU puffins and mean T

puffin_temp <- merge(eu_puffins, mean_t_data)

# Select values to correlate
corpuffin_temp <- puffin_temp %>%
  select(id, year, mean_tmax, mean_tmin, pop_trend) %>%
  spread(key = "id", value = "pop_trend") %>%
  drop_na()

# Correlation

cormat <- round(cor(corpuffin_temp[, c(2:9)]),2) %>%
  melt()
```

The figures are presented and the code for them is run within the chunks
(but is not made visible).

## Figure 1.
```{r echo=F, out.width="100%", fig.show='hold', fig.pos='H', fig.width=7, fig.height=5, fig.cap="Comparison of EU puffins population trends with average max temperature (째C)."}

# Compare Puffin populations with temperature

  # Max temps
ggplot(puffin_temp) +
  ggtitle("Comparison between puffin abundance and max. temperature changes (1930-2018)\n") +
  theme(plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 11, face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.minor = element_line(colour = "grey", size = 0.1 ),
        panel.border = element_rect(colour = "grey", fill = NA)) +
  scale_colour_discrete(name = "Puffins i.d.") +
  xlab("\nMean max. temperature (째C)") + ylab("Population trend\n") +
  geom_point(aes(x = mean_tmax, y = pop_trend,
                 colour = as.factor(id))) +
  geom_smooth(aes(x = mean_tmax, y = pop_trend,
                  colour = as.factor(id)),
              method = 'lm')
```

## Figure 2
```{r echo=F, out.width="100%", fig.show='hold', fig.pos='H', fig.width=7, fig.height=5, fig.cap="Comparison of EU puffins population trends with average min temperature (째C)."}

# Compare Puffin populations with temperature

  # Min temps
ggplot(puffin_temp) +
  ggtitle("Comparison between puffin abundance and min. temperature changes (1930-2018)\n") +
  theme(plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 11, face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.minor = element_line(colour = "grey", size = 0.1 ),
        panel.border = element_rect(colour = "grey", fill = NA)) +
  scale_colour_discrete(name = "Puffins i.d.") +
  xlab("\nMean min. temperature (째C)") + ylab("Population trend\n") +
  geom_point(aes(x = mean_tmin, y = pop_trend,
                 colour = as.factor(id))) +
  geom_smooth(aes(x = mean_tmin, y = pop_trend,
                  colour = as.factor(id)),
              method = 'lm')
```

## Figure 3.

```{r echo=F, out.width="100%", fig.show='hold', fig.pos='H', fig.width=7, fig.height=5, fig.cap="Correlation mat between years and puffin abundance/mean temperatures"}

# Correlation mat

ggplot(cormat,
       aes(x = Var1, y = Var2, fill = value)) +
  ggtitle("Correlation mat of population trends\n") +
  theme(plot.title = element_text(hjust = 0.5, vjust = 0.5,
                                  size = 11, face = "bold")) +
  geom_tile() +
  xlab(" ") + ylab(" ")
```


\newpage

# References


A list of references here. 


\newpage


# Appendix

This part is to add the appendix.
